Evaluating Toronto's Fire Incidents to Extrapolate Future Occurence Data

### 2024-04-02 paper/paper.qmd
updated introduction section of paper

Fires pose a significant threat to both residential and commercial properties, often resulting in substantial financial losses and, more importantly, endangering lives. Developing a better understanding of the factors that influence the amount in damages per fire incident is crucial for implementing effective fire prevention measures to minimize the impact of such incidents. In this analysis, we delve into the extensive dataset provided by Toronto Open Data [@citeTorontoFireIncidents] which captures fire incidents within the Toronto area as defined by Ontario Fire Marshal (OFM) from 2011-2022. Through this dataset, we aim to predict the estimated dollar loss per fire incident across different property uses and identify the factors contributing to these losses.

The motivation behind this study stems from the necessity to comprehend the underlying patterns and determinants of fire incident costs. By dissecting the data provided by Toronto Open Data, we seek to not only predict the estimated dollar loss but also to identify significant factors influencing these losses. Our analysis extends beyond mere prediction, delving into factors such as Toronto Fire Services' response time, area of origin, and the presence and functionality of fire alarms. These additional variables are essential in understanding the dynamics of fire incidents and formulating informed strategies for fire prevention and mitigation.

Despite the abundance of data available, there remains a gap in comprehensively analyzing the factors contributing to the estimated dollar loss per fire incident in the Toronto area. To address this gap, we employ a data-driven approach, utilizing statistical modeling techniques to uncover patterns and relationships within the dataset. Our methodology involves exploring various factors such as property use, possible causes of the incident (e.g., mechanical/electrical failures), and the presence and functionality of fire alarms. Additionally, we incorporate the number of responding personnel from the Toronto Fire Services (TFS) as a variable to assess its impact on the estimated dollar loss.

The findings of this analysis hold significant implications for fire safety practices and policies in the Toronto area. By understanding the factors influencing the estimated dollar loss per fire incident, stakeholders can prioritize resources and implement targeted interventions to mitigate fire risks effectively. This paper is structured as follows: we first provide an overview of the dataset and variables considered, followed by the methodology employed for analysis. Subsequently, we present our findings, discussing the relationships uncovered and their implications for fire safety. Finally, we conclude by highlighting the importance of our findings and suggesting avenues for future research in this domain.

>>>>

Fires pose a significant threat to both residential and commercial properties, often resulting in substantial financial losses and, more importantly, endangering lives. Developing a better understanding of the factors that influence the amount in damages per fire incident is crucial for implementing effective fire prevention measures to minimize the impact of such incidents. In this analysis, we delve into the extensive dataset provided by Toronto Open Data [@citeTorontoFireIncidents] which captures fire incidents within the Toronto area as defined by Ontario Fire Marshal (OFM) from 2011 to 2022. Through this dataset, we aim to predict the estimated dollar loss per fire incident across different property uses and identify the factors contributing to these losses.

The motivation behind this analysis is that fire safety matters and is important. Beyond the protection of assets, it's about keeping people safe. In order to ensure the safety of fellow Canadians, it is imperative to comprehend the underlying patterns and determinants of fire incident. By dissecting the data provided by Toronto Open Data, we seek to not only predict the estimated amount in damages but also to identify significant factors influencing these losses. Analyzing factors such as Toronto Fire Services' response time, area of origin, and the presence, functionality, and configuration of fire alarms. These additional variables are essential in understanding the dynamics of fire incidents and formulating informed strategies for fire prevention and mitigation.

Despite the abundance of data available, there remains a gap in comprehensively analyzing the factors contributing to the estimated dollar loss per fire incident in the Toronto area. To address this gap, we employ a data-driven approach, utilizing statistical modeling techniques to uncover patterns and relationships within the dataset. Our methodology involves exploring various factors such as property use, possible causes of the incident (e.g., mechanical/electrical failures), and the presence of functioning fire alarms. Additionally, we incorporate the number of responding personnel from the Toronto Fire Services (TFS) as a variable to assess its impact on the estimated dollar loss.

The findings of this analysis hold significant implications for fire safety practices and policies in the Toronto area. By understanding the factors influencing the estimated dollar loss per fire incident, stakeholders can prioritize resources and allocate them accordingly to mitigate fire risks effectively. This paper is structured as follows: we first provide an overview of the dataset and variables of interest, followed by the methodology employed for analysis in @sec-data. Subsequently, we present our findings, discussing the relationships uncovered and their implications for fire safety in @sec-results. Finally, we conclude by highlighting the importance of our findings and suggesting avenues for future research in this domain concerning fire safety in @sec-discussion.
